
# **RNA Data preprocessing**
## Loading packages

Load the IOBR package in your R session after the installation is complete:
```{r, eval=TRUE, warning=FALSE, message=FALSE}
library(IOBR)
library(tidyverse)
library(clusterProfiler)
```

## Download array data using `GEOquery`
Obtaining data set from GEO [Gastric cancer: GSE62254](https://pubmed.ncbi.nlm.nih.gov/25894828/) using `GEOquery` R package.
```{r,message=FALSE,warning=FALSE}
if (!requireNamespace("GEOquery", quietly = TRUE))  BiocManager::install("GEOquery")
library("GEOquery")
# NOTE: This process may take a few minutes which depends on the internet connection speed. Please wait for its completion.
eset_geo<-getGEO(GEO     = "GSE62254", getGPL  = F, destdir = "./")
eset    <-eset_geo[[1]]
eset    <-exprs(eset)
eset[1:5,1:5]
```

## Gene Annotation

Annotation of genes in the expression matrix and removal of duplicate genes.
```{r,message=FALSE,warning=FALSE}
# Load the annotation file `anno_hug133plus2` in IOBR.
head(anno_hug133plus2)

```

```{r,message=FALSE,warning=FALSE}
# Load the annotation file `anno_grch38` in IOBR.
head(anno_grch38)

```

```{r,message=FALSE,warning=FALSE}
# Load the annotation file `anno_gc_vm32` in IOBR for mouse RNAseq data
head(anno_gc_vm32)

```

### For Array data: HGU133PLUS-2 (Affaymetrix)
```{r,message=FALSE,warning=FALSE}

# Conduct gene annotation using `anno_hug133plus2` file; If identical gene symbols exists, these genes would be ordered by the mean expression levels. The gene symbol with highest mean expression level is selected and remove others. 

eset<-anno_eset(eset       = eset,
                annotation = anno_hug133plus2,
                symbol     = "symbol",
                probe      = "probe_id",
                method     = "mean")
eset[1:5, 1:3]
```

## Download RNAseq data using `UCSCXenaTools`

```{r, eval=FALSE}
if (!requireNamespace("UCSCXenaTools", quietly = TRUE))   BiocManager::install("UCSCXenaTools")
library(UCSCXenaTools)
# NOTE: This process may take a few minutes which depends on the internet connection speed. Please wait for its completion.
eset_stad<-XenaGenerate(subset = XenaCohorts =="GDC TCGA Stomach Cancer (STAD)") %>% 
  XenaFilter(filterDatasets    = "TCGA-STAD.htseq_counts.tsv") %>% 
  XenaQuery() %>%
  XenaDownload() %>% 
  XenaPrepare()
eset_stad[1:5, 1:3]
```

## Normalization and Gene annotation
Transform gene expression matrix into TPM format, and conduct subsequent annotation. 
```{r, eval=FALSE}

# Remove the version numbers in Ensembl ID.
eset_stad$Ensembl_ID<-substring(eset_stad$Ensembl_ID, 1, 15)
eset_stad<-column_to_rownames(eset_stad, var = "Ensembl_ID")

# Revert back to original format because the data from UCSC was log2(x+1)transformed.
eset_stad<-(2^eset_stad)+1

eset_stad<-count2tpm(countMat = eset_stad, idType = "Ensembl", org="hsa", source = "local" )

eset_stad[1:5,1:5]
```

## Identifying outlier samples

Take ACRG microarray data for example
```{r, message = F, warning = F,  fig.width= 7, fig.height=7, fig.align='center'}

res <- find_outlier_samples(eset = eset, project = "ACRG", show_plot = TRUE)
```

Removing potential outlier samples
```{r, message = F, }
eset1 <- eset[, !colnames(eset)%in%res]
```

## PCA analysis of molecular subtypes

```{r, message = F, warning = F, fig.width= 7.6, fig.height=7, fig.align='center'}
data("pdata_acrg")
res<- iobr_pca(data       = eset1,
              is.matrix   = TRUE,
              scale       = TRUE,
              is.log      = FALSE,
              pdata       = pdata_acrg, 
              id_pdata    = "ID", 
              group       = "Subtype",
              geom.ind    = "point", 
              cols        = "normal",
              palette     = "jama", 
              repel       = FALSE,
              ncp         = 5,
              axes        = c(1, 2),
              addEllipses = TRUE)
res
```


## Batch effect correction

### For microarray data
Obtaining another data set from GEO [Gastric cancer: GSE57303](https://www.ncbi.nlm.nih.gov/pubmed/24935174/) using `GEOquery` R package.
```{r,message=FALSE,warning=FALSE}
# NOTE: This process may take a few minutes which depends on the internet connection speed. Please wait for its completion.
eset_geo<-getGEO(GEO     = "GSE57303", getGPL  = F, destdir = "./")
eset2    <-eset_geo[[1]]
eset2    <-exprs(eset2)
eset2[1:5,1:5]
```

Annotation of genes in the expression matrix and removal of duplicate genes.
```{r,message=FALSE,warning=FALSE}
eset2<-anno_eset(eset       = eset2,
                 annotation = anno_hug133plus2,
                 symbol     = "symbol",
                 probe      = "probe_id",
                 method     = "mean")
eset2[1:5, 1:5]
```


```{r, message=FALSE, fig.width= 11, fig.height = 5, fig.align='center'}
eset_com <- remove_batcheffect( eset1       = eset1,  
                                eset2       = eset2,   
                                eset3       = NULL,
                                id_type     = "symbol",
                                data_type   = "array", 
                                cols        = "normal", 
                                palette     = "jama", 
                                log2        = TRUE, 
                                check_eset  = TRUE,
                                adjust_eset = TRUE,
                                repel       = FALSE,
                                path        = "result")
dim(eset_com)
```


### For RNAseq count data

```{r, message=FALSE, fig.width= 11, fig.height = 5, fig.align='center'}

data("eset_stad", package = "IOBR")
head(eset_stad)

data("eset_blca", package = "IOBR")
head(eset_blca)

eset_com <- remove_batcheffect(eset_stad, eset_blca, id_type = "ensembl", data_type = "count")

# The returned matrix is the count matrix after removing the batches.
head(eset_com)
```

## References

Wang et al., (2019). The UCSCXenaTools R package: a toolkit for accessing genomics data from UCSC Xena platform, from cancer multi-omics to single-cell RNA-seq. Journal of Open Source Software, 4(40), 1627, https://doi.org/10.21105/joss.01627

Zhang et al., ComBat-seq: batch effect adjustment for RNA-seq count data, NAR Genomics and Bioinformatics, Volume 2, Issue 3, September 2020, lqaa078, https://doi.org/10.1093/nargab/lqaa078

Leek, J. T., et al., (2012). The sva package for removing batch effects and other unwanted variation in high-throughput experiments. Bioinformatics, 28(6), 882-883.


