

# **Signature Score and Relevant phenotypes**

## Loading packages
Load the IOBR package in your R session after the installation is complete:
```{r, eval=TRUE, warning=FALSE, message=FALSE}
library(IOBR)
library(survminer)
library(tidyverse)
```

## Downloading data for example
Obtaining data set from GEO [Gastric cancer: GSE62254](https://pubmed.ncbi.nlm.nih.gov/25894828/) using `GEOquery` R package.
```{r,message=FALSE,warning=FALSE}
if (!requireNamespace("GEOquery", quietly = TRUE))  BiocManager::install("GEOquery")
library("GEOquery")
# NOTE: This process may take a few minutes which depends on the internet connection speed. Please wait for its completion.
eset_geo <- getGEO(GEO = "GSE62254", getGPL  = F, destdir = "./")
eset    <- eset_geo[[1]]
eset    <- exprs(eset)
eset[1:5,1:5]
```

## Gene Annotation
Annotation of genes in the expression matrix and removal of duplicate genes.
```{r,message=FALSE,warning=FALSE}

# Load the annotation file `anno_hug133plus2` in IOBR.
head(anno_hug133plus2)
# Conduct gene annotation using `anno_hug133plus2` file; If identical gene symbols exists, these genes would be ordered by the mean expression levels. The gene symbol with highest mean expression level is selected and remove others. 

eset<-anno_eset(eset       = eset,
                annotation = anno_hug133plus2,
                symbol     = "symbol",
                probe      = "probe_id",
                method     = "mean")
eset[1:5, 1:3]
```


## Estimation of signatures
```{r, message = F, warning = F}

sig_tme<-calculate_sig_score(pdata           = NULL,
                             eset            = eset,
                             signature       = signature_collection,
                             method          = "pca",
                             mini_gene_count = 2)

sig_tme <- t(column_to_rownames(sig_tme, var = "ID"))
sig_tme[1:5, 1:3]
```

## Combining score data and phenotype data

```{r, message = F, warning = F, fig.width= 7, fig.height=7.5, fig.align='center'}
data("pdata_acrg", package = "IOBR")
head(pdata_acrg)
input <- combine_pd_eset(eset = sig_tme, pdata = pdata_acrg, scale = T)
```


## Identifying features associated with survival

```{r, message = F, warning = F, fig.width= 7, fig.height=7.5, fig.align='center'}
res<- batch_surv(pdata    = input,
                 time     = "OS_time", 
                 status   = "OS_status", 
                 variable = colnames(input)[69:ncol(input)])
head(res)
```

Use forest plots `sig_forest` to show the most relevant variables to overall survival
```{r, message = F, warning = F, fig.width= 7, fig.height=7.5, fig.align='center'}
res<- res[nchar(res$ID)<=28, ]
p1<- sig_forest(res, signature = "ID", n = 20)
```

## Visulization using heatmap

Relationship between Signatures and molecular typing.
Heatmap visualisation using `IOBR`'s `sig_heatmap`
```{r, message=FALSE, fig.width= 9, fig.height = 6, fig.align='center'}
p2 <- sig_heatmap(input         = input, 
                  features      = res$ID[1:20],
                  group         = "Subtype", 
                  palette_group = "jama", 
                  palette       = 6,
                  path          = "result" )
```

## Focus on target signatures
```{r, fig.show= 'hide', fig.width= 12, fig.height=8.5, message = F, warning = F}
p1 <- sig_box(data           = input, 
              signature      = "Glycogen_Biosynthesis",
              variable       = "Subtype",
              jitter         = FALSE,
              cols           =  NULL,
              palette        = "jama",
              show_pvalue    = TRUE,
              size_of_pvalue = 5,
              hjust          = 1, 
              angle_x_text   = 60, 
              size_of_font   = 8)

p2 <- sig_box(data           = input, 
              signature      = "Pan_F_TBRs",
              variable       = "Subtype",
              jitter         = FALSE,
              cols           = NULL,
              palette        = "jama",
              show_pvalue    = TRUE,
              angle_x_text   = 60, 
              hjust          = 1, 
              size_of_pvalue = 5, 
              size_of_font   = 8)

p3 <- sig_box(data           = input, 
              signature      = "Immune_Checkpoint",
              variable       = "Subtype",
              jitter         = FALSE,
              cols           = NULL,
              palette        = "jama",
              show_pvalue    = TRUE,
              angle_x_text   = 60, 
              hjust          = 1, 
              size_of_pvalue = 5, 
              size_of_font   = 8)
```


```{r, message = F, warning = F, fig.width= 10, fig.height=6, fig.align='center'}
p1|p2|p3
```


## Survival analysis and visulization

### Kaplan-Meier plot
Displaying the outcomes of survival analyses using Kaplan-Meier plot. Multiple stratifications of the signature were used to judge the efficacy of this metric in predicting patient survival.
```{r, message = F, warning = F, fig.width= 16, fig.height=8}
res <-       sig_surv_plot(input_pdata       = input, 
                           signature         = "Glycogen_Biosynthesis",
                           cols              = NULL, 
                           palette           = "jama",
                           project           = "ACRG",
                           time              = "OS_time",
                           status            = "OS_status",
                           time_type         = "month",
                           save_path         = "result")
res$plots
```

### Time-Dependent ROC curve

```{r, message = F, warning = F, fig.width= 14, fig.height=6.5}
p1<- roc_time(input      = input,  
             vars       = "Glycogen_Biosynthesis", 
             time       = "OS_time",
             status     = "OS_status", 
             time_point = c(12, 24, 36), 
             time_type  = "month",
             palette    = "jama",
             cols       = "normal",
             seed       = 1234, 
             show_col   = FALSE, 
             path       = "result", 
             main       = "OS",
             index      = 1,
             fig.type   = "pdf",
             width      = 5,
             height     = 5.2)

p2<- roc_time(input      = input,  
             vars       = "Glycogen_Biosynthesis", 
             time       = "RFS_time",
             status     = "RFS_status", 
             time_point = c(12, 24, 36), 
             time_type  = "month",
             palette    = "jama",
             cols       = "normal",
             seed       = 1234, 
             show_col   = FALSE, 
             path       = "result", 
             main       = "OS",
             index      = 1,
             fig.type   = "pdf",
             width      = 5,
             height     = 5.2)
p1|p2
```


## Batch correlation analysis 

### Finding continuity variables associated with signatures
Identifying genes or signatures related to the target signatures

#### Correlation between two variables
```{r}
res <- batch_cor(data = input, target = "Glycogen_Biosynthesis", feature = colnames(input)[69:ncol(input)])
head(res)
```

```{r, fig.show= 'hide', fig.width= 12, fig.height=9, message = F, warning = F}
p1<- get_cor(eset = sig_tme, pdata = pdata_acrg, is.matrix = TRUE, var1 = "Glycogen_Biosynthesis", 
             var2 = "TMEscore_CIR", subtype = "Subtype", palette = "aaas", path = "result")
p2<- get_cor(eset = sig_tme, pdata = pdata_acrg, is.matrix = TRUE, var1 = "Glycogen_Biosynthesis", 
             var2 = "TGFβ_myCAF", subtype = "Subtype", palette = "aaas", path = "result")
```

```{r, fig.width= 12, fig.height=6, message = F, warning = F}
p1|p2
```

#### Demonstrate correlation between multiple variables

Visualisation via correlation matrix
```{r, fig.width= 10, fig.height=4, message = F, warning = F}
feas1 <- c("Glycogen_Biosynthesis", "Ferroptosis")
feas2 <- c("Glutathione_Metabolism", "TMEscore_CIR", "Purine_Metabolism", "ICB_resistance_Peng_et_al", "Interleukins_Li_et_al", "TLS_Nature")
p <- get_cor_matrix(data           = input, 
                    feas1          = feas2, 
                    feas2          = feas1,
                    method         = "pearson",
                    font.size.star = 8, 
                    font.size      = 15, 
                    fill_by_cor    = FALSE, 
                    round.num      = 1, 
                    path           = "result")
```

Demonstrate the correlation between signatures and genes
```{r, fig.width= 11, fig.height=4, message = F, warning = F}
input2 <- combine_pd_eset(eset = eset, pdata =  input[, c("ID", "Glycogen_Biosynthesis", "TLS_Nature", "Ferroptosis")])
feas1 <- c("Glycogen_Biosynthesis","TLS_Nature", "Ferroptosis")
feas2 <- signature_collection$CD_8_T_effector
feas2
p <- get_cor_matrix(data           = input2, 
                    feas1          = feas2, 
                    feas2          = feas1,
                    method         = "pearson",
                    scale          = T, 
                    font.size.star = 8, 
                    font.size      = 15, 
                    fill_by_cor    = FALSE, 
                    round.num      = 1,
                    path           = "result")
```

Users can customize the image using parameters.
```{r, fig.width= 11, fig.height=4, message = F, warning = F}
p <- get_cor_matrix(data           = input2, 
                    feas1          = feas2, 
                    feas2          = feas1,
                    method         = "pearson",
                    scale          = T, 
                    font.size.star = 8, 
                    font.size      = 15, 
                    fill_by_cor    = TRUE, 
                    round.num      = 2,
                    path           = "result")
```


### Identifying Category Variables Linked to Signatures

#### For binary variable
```{r, fig.width= 11, fig.height=4, message = F, warning = F}

res <- batch_wilcoxon(data = input, target = "TMEscore_binary", feature = colnames(input)[69:ncol(input)])
head(res)
```
```{r, fig.show= 'hide', fig.width= 12, fig.height=8.5, message = F, warning = F}
p1 <- sig_box(data           = input, 
              signature      = res$sig_names[1],
              variable       = "TMEscore_binary",
              jitter         = FALSE,
              cols           =  NULL,
              palette        = "jco",
              show_pvalue    = TRUE,
              size_of_pvalue = 5,
              hjust          = 1, 
              angle_x_text   = 60, 
              size_of_font   = 8)

p2 <- sig_box(data           = input, 
              signature      = res$sig_names[2],
              variable       = "TMEscore_binary",
              jitter         = FALSE,
              cols           = NULL,
              palette        = "jco",
              show_pvalue    = TRUE,
              angle_x_text   = 60, 
              hjust          = 1, 
              size_of_pvalue = 5, 
              size_of_font   = 8)

p3 <- sig_box(data           = input, 
              signature      = res$sig_names[3],
              variable       = "TMEscore_binary",
              jitter         = FALSE,
              cols           = NULL,
              palette        = "jco",
              show_pvalue    = TRUE,
              angle_x_text   = 60, 
              hjust          = 1, 
              size_of_pvalue = 5, 
              size_of_font   = 8)
```


```{r, message = F, warning = F, fig.width= 9, fig.height=5, fig.align='center'}
p1|p2|p3
```


### For multicategorical variables (>2 subgroups)
```{r, fig.width= 11, fig.height=4, message = F, warning = F}

res <- batch_kruskal(data = input, group = "Subtype", feature = colnames(input)[69:ncol(input)])
head(res)
```
```{r, fig.show= 'hide', fig.width= 12, fig.height=8.5, message = F, warning = F}
p1 <- sig_box(data           = input, 
              signature      = res$sig_names[1],
              variable       = "Subtype",
              jitter         = FALSE,
              cols           =  NULL,
              palette        = "jco",
              show_pvalue    = TRUE,
              size_of_pvalue = 5,
              hjust          = 1, 
              angle_x_text   = 60, 
              size_of_font   = 8)

p2 <- sig_box(data           = input, 
              signature      = res$sig_names[2],
              variable       = "Subtype",
              jitter         = FALSE,
              cols           = NULL,
              palette        = "jco",
              show_pvalue    = TRUE,
              angle_x_text   = 60, 
              hjust          = 1, 
              size_of_pvalue = 5, 
              size_of_font   = 8)

p3 <- sig_box(data           = input, 
              signature      = res$sig_names[3],
              variable       = "Subtype",
              jitter         = FALSE,
              cols           = NULL,
              palette        = "jco",
              show_pvalue    = TRUE,
              angle_x_text   = 60, 
              hjust          = 1, 
              size_of_pvalue = 5, 
              size_of_font   = 8)
```


```{r, message = F, warning = F, fig.width= 10, fig.height=6, fig.align='center'}
p1|p2|p3
```

## Reference

Cristescu, R., Lee, J., Nebozhyn, M. et al. Molecular analysis of gastric cancer identifies subtypes associated with distinct clinical outcomes. Nat Med 21, 449–456 (2015). https://doi.org/10.1038/nm.3850

Dongqiang Zeng, ..., WJ Liao et al., Tumor Microenvironment Characterization in Gastric Cancer Identifies Prognostic and Immunotherapeutically Relevant Gene Signatures, Cancer Immunol Res (2019) 7 (5): 737–750. https://doi.org/10.1158/2326-6066.CIR-18-0436
