
# **TME Interaction analysis**

## Loading packages

```{r, eval=TRUE, warning=FALSE, message=FALSE}
library(IOBR)
```

## Downloading data for example
Obtaining data set from GEO [Gastric cancer: GSE62254](https://pubmed.ncbi.nlm.nih.gov/25894828/) using `GEOquery` R package.
```{r,message=FALSE,warning=FALSE}
if (!requireNamespace("GEOquery", quietly = TRUE))  BiocManager::install("GEOquery")
library("GEOquery")
# NOTE: This process may take a few minutes which depends on the internet connection speed. Please wait for its completion.
eset_geo<- getGEO(GEO     = "GSE62254", getGPL  = F, destdir = "./")
eset    <- eset_geo[[1]]
eset    <- exprs(eset)
eset[1:5,1:5]
```

## Gene Annotation: HGU133PLUS-2 (Affaymetrix)
```{r,message=FALSE,warning=FALSE}

# Conduct gene annotation using `anno_hug133plus2` file; If identical gene symbols exists, these genes would be ordered by the mean expression levels. The gene symbol with highest mean expression level is selected and remove others. 

eset<-anno_eset(eset       = eset,
                annotation = anno_hug133plus2,
                symbol     = "symbol",
                probe      = "probe_id",
                method     = "mean")
eset[1:5, 1:3]
```

## TME deconvolution using CIBERSORT algorithm
```{r, message=FALSE, warning=FALSE}
cell <- deconvo_tme(eset = eset, method = "cibersort", arrays = TRUE, perm = 500, absolute.mode = TRUE)
head(cell)
```
## Identifying TME patterns
Identification of optimal clustering based on cellular infiltration patterns in the microenvironment.

```{r, message=FALSE, warning=FALSE}
tme <- tme_cluster(input = cell, features = colnames(cell)[2:23], id = "ID", scale = TRUE, method = "kmeans", max.nc = 5)
```

Use of heatmaps to reflect cellular differences between TME subtypes
```{r, message = F, warning = F, fig.width=6, fig.height=4}
colnames(tme) <- gsub(colnames(tme), pattern = "_CIBERSORT", replacement = "")
res <- sig_heatmap(input = tme, features = colnames(tme)[3:ncol(tme)], group = "cluster", path = "result", palette = 6)
```

## Cell abundance of each cluster

```{r,fig.show= 'hide', message=FALSE, warning=FALSE}
cols <- c('#2692a4','#fc0d3a','#ffbe0b')
p1 <- sig_box(tme, variable = "cluster", signature = "Macrophages_M1", jitter = TRUE,
              cols =  cols, show_pvalue = TRUE, size_of_pvalue = 4)

p2 <- sig_box(tme, variable = "cluster", signature = "Mast_cells_activated", 
              jitter = TRUE, cols =  cols, show_pvalue = TRUE, size_of_pvalue = 4)

p3 <- sig_box(tme, variable = "cluster", signature = "Macrophages_M2", 
              jitter = TRUE, cols =  cols, show_pvalue = TRUE, size_of_pvalue = 4)
```

```{r,fig.width= 11, fig.height=4.5}
p1|p2|p3
```

## DEG analysis between TME subtypes

Identifying TME subtypes related differential genes using `find_markers_in_bulk`.

We have developed a reliable classifier for the tumour microenvironment in gastric cancer using the same analysis pipeline[TMEclassifier](https://github.com/LiaoWJLab/TMEclassifier). The classifier was constructed by identifying the most robust gastric cancer TME classification through parsing the tumour microenvironment using the `tme_cluster` method. Next, genes specifically expressed by each microenvironmental subtype are obtained using the `find_markers_in_bulk method`. Finally, a machine learning approach was used to construct the classifier model.

```{r, message = F, warning = F}
library(Seurat)
res <- find_markers_in_bulk(pdata      = tme, 
                            eset       = eset, 
                            group      = "cluster", 
                            nfeatures  = 2000, 
                            top_n      = 50, 
                            thresh.use = 0.15, 
                            only.pos   = TRUE, 
                            min.pct    = 0.10)
top15 <-  res$top_markers %>% dplyr:: group_by(cluster) %>%  dplyr::top_n(15, avg_log2FC)
top15$gene
```

Heatmap visualisation using `Seurat`'s `DoHeatmap`
```{r,fig.show= 'hide',message=FALSE}
#Defining cluster colors
cols <- c('#2692a4','#fc0d3a','#ffbe0b')
p1 <- DoHeatmap(res$sce, top15$gene, group.colors = cols )+
  scale_fill_gradientn(colours = rev(colorRampPalette(RColorBrewer::brewer.pal(11,"RdBu"))(256)))

```

Extracting variables from the expression matrix to merge with TME subtype
```{r,fig.show= 'hide', message=FALSE}

input <- combine_pd_eset(eset = eset, pdata = tme, feas = top15$gene, scale = T)
p2 <- sig_box(input, variable = "cluster", signature = "IFNG", jitter = TRUE,
              cols =  cols, show_pvalue = TRUE, size_of_pvalue = 4)

p3 <- sig_box(input, variable = "cluster", signature = "IL1A", 
              jitter = TRUE, cols =  cols, show_pvalue = TRUE, size_of_pvalue = 4)
```

Combining the results obtained above
```{r, fig.width= 12, fig.height=9, message = F, warning = F}

# if (!requireNamespace("patchwork", quietly = TRUE))   install.packages("patchwork")
library(patchwork)
p <- (p1|p2/p3) + plot_layout(widths = c(2.3,1))
p + plot_annotation(tag_levels = 'A')
```

## Identifying LR associated with TME clusters

The `LR_cal` method originates from the study:
**Lapuente-Santana, Ó., van Genderen, M., Hilbers, P., Finotello, F., & Eduati, F. (2021). "Interpretable systems biomarkers predict response to immune-checkpoint inhibitors."** Patterns (New York, N.Y.), 2(8), 100293. [DOI: 10.1016/j.patter.2021.100293](https://doi.org/10.1016/j.patter.2021.100293).
The method calculates interaction weights for 813 ligand-receptor (LR) pairs. For this function to run successfully, you need to install the `easier` package following the tutorial available at [easier GitHub repository](https://github.com/olapuentesantana/easier).

```{r, message = F, warning = F}
lr_data <- LR_cal(eset = eset, data_type = "tpm", id_type = "symbol")

```


```{r, message = F, warning = F}
lr <- column_to_rownames(lr_data,var = "ID") %>% t() %>% as.data.frame()
res <- find_markers_in_bulk(pdata      = tme, 
                            eset       = lr, 
                            group      = "cluster", 
                            nfeatures  = 2000, 
                            top_n      = 50, 
                            thresh.use = 0.15, 
                            only.pos   = TRUE, 
                            min.pct    = 0.10)
top15 <-  res$top_markers %>% dplyr:: group_by(cluster) %>%  dplyr::top_n(15, avg_log2FC)
top15$gene
```

Heatmap visualisation using `Seurat`'s `DoHeatmap`
```{r}
#cols <- c('#2692a4','#fc0d3a','#ffbe0b')
p1 <- DoHeatmap(res$sce, top15$gene, group.colors = cols )+
  scale_fill_gradientn(colours = rev(colorRampPalette(RColorBrewer::brewer.pal(11,"RdBu"))(256)))
```

Extracting variables from the expression matrix to merge with TME subtype
```{r,fig.show= 'hide', message=FALSE}
top15$gene  <- gsub(top15$gene, pattern = "-", replacement = "\\_")
input <- combine_pd_eset(eset = lr, pdata = tme, feas = top15$gene, scale = T)

p2 <- sig_box(input, variable = "cluster", signature = top15$gene[1], jitter = TRUE,
              cols =  cols, show_pvalue = TRUE, size_of_pvalue = 4)

p3 <- sig_box(input, variable = "cluster", signature = top15$gene[5], 
              jitter = TRUE, cols =  cols, show_pvalue = TRUE, size_of_pvalue = 4)
```

Combining the results obtained above
```{r, fig.width= 12, fig.height=9, message = F, warning = F}

# if (!requireNamespace("patchwork", quietly = TRUE))   install.packages("patchwork")
library(patchwork)
p <- (p1|p2/p3) + plot_layout(widths = c(2.3,1))
p + plot_annotation(tag_levels = 'A')
```

## Identifying signatures associated with TME clusters

Calculate TME associated signatures-(through PCA method).
```{r, message = F, warning = F}
sig_tme<-calculate_sig_score(pdata           = NULL,
                             eset            = eset,
                             signature       = signature_collection,
                             method          = "pca",
                             mini_gene_count = 2)
sig_tme <- t(column_to_rownames(sig_tme, var = "ID"))
sig_tme[1:5, 1:3]
```

Finding characteristic variables associated with TME clusters
```{r, message = F, warning = F}
res <- find_markers_in_bulk(pdata = tme, eset = sig_tme, group = "cluster", nfeatures = 1000, top_n = 20, min.pct = 0.10)

top15 <-  res$top_markers %>% dplyr:: group_by(cluster) %>%  dplyr::top_n(15, avg_log2FC)

p1 <- DoHeatmap(res$sce, top15$gene, group.colors = cols)+
  scale_fill_gradientn(colours = rev(colorRampPalette(RColorBrewer::brewer.pal(11,"RdBu"))(256)))
```

Visualizing results and selecting feature variables
```{r,fig.show= 'hide', message=FALSE}
top15$gene  <- gsub(top15$gene, pattern = "-", replacement = "\\_")
input <- combine_pd_eset(eset = sig_tme, pdata = tme, feas = top15$gene, scale = T)

p2 <- sig_box(input, variable = "cluster", signature = "CD_8_T_effector", jitter = TRUE,
              cols =  cols, show_pvalue = TRUE, size_of_pvalue = 4, size_of_font = 6)

p3 <- sig_box(input, variable = "cluster", signature = "Neutrophils_Bindea_et_al",  
              jitter = TRUE, cols =  cols, show_pvalue = TRUE, size_of_pvalue = 4, size_of_font = 6)
```


```{r, fig.width= 12, fig.height=9}
p <- (p1|p2/p3) + plot_layout(widths = c(2.3,1))
p + plot_annotation(tag_levels = 'A')
```


Survival differences between tumour microenvironment subtypes

```{r, fig.width= 5, fig.height=8}
library(survminer)
data(pdata_acrg, package = "IOBR")
input <- merge(pdata_acrg, input, by = "ID")
p1<-surv_group(input_pdata       = input,
               target_group      = "cluster",
               ID                = "ID",
               reference_group   = "High",
               project           = "ACRG",
               cols              = cols, 
               time              = "OS_time",
               status            = "OS_status",
               time_type         = "month",
               save_path         = "result")
p1
```

Relationship between tumour microenvironmental subtypes and other subtypes

```{r,fig.show= 'hide',message=FALSE}
p1<- percent_bar_plot(input, x = "cluster" , y = "Subtype", palette = "jama", axis_angle = 60)
p2<- percent_bar_plot(input, x = "cluster" , y = "Lauren", palette = "jama", axis_angle = 60)
p3<- percent_bar_plot(input, x = "cluster" , y = "TMEscore_binary", palette = "jama", axis_angle = 60)
```

```{r,fig.width= 12, fig.height=6}
p1|p2|p3
```

## References

Cristescu, R., Lee, J., Nebozhyn, M. et al. Molecular analysis of gastric cancer identifies subtypes associated with distinct clinical outcomes. Nat Med 21, 449–456 (2015). https://doi.org/10.1038/nm.3850

[CIBERSORT](https://cibersort.stanford.edu/); Newman, A. M., Liu, C. L., Green, M. R., Gentles, A. J., Feng, W., Xu, Y., … Alizadeh, A. A. (2015). Robust enumeration of cell subsets from tissue expression profiles. Nature Methods, 12(5), 453–457.  https://doi.org/10.1038/nmeth.3337; 

Seurat: Hao and Hao et al. Integrated analysis of multimodal single-cell data. Cell (2021)

[easier](https://doi.org/10.1016/j.patter.2021.100293); Lapuente-Santana, Ó., van Genderen, M., Hilbers, P., Finotello, F., & Eduati, F. (2021). 'Interpretable systems biomarkers predict response to immune-checkpoint inhibitors.' Patterns (New York, N.Y.), 2(8), 100293.

